Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta

### 2024-07-31 response/response.tex

\section{Responses to the Reviewers}

\subsection{Comments from Reviewer 1}

\begin{itemize}

\item Reviewer 1 wrote:
\begin{quote}
``No Verification \& Validation of the simulation model is described. At least some discussion regarding the validity of the model has to be included in the case study.''
\end{quote}
Thorough verification of the models is done by the automated code testing suits provided in the code base, a paragraph on this is now included under Section 5.5. Validation of this model was difficult, and relied on the expert opinion of our collaborators ar 118 and 119. This is now discussed in Section 6.

@@

\end{itemize}

\subsection{Comments from Reviewer 2}

\begin{itemize}

\item Reviewer 2 wrote:
\begin{quote}
``The introduction makes a case for using survival vs reponse time as a measure for ambulance allocation (although this is not the only measure in the results). What would this do to fairness in the allocations? Would far away patients with serious conditions (ie. Needing a fast response time) just be discarded as it is expensive to cater for them (in terms of overall survival of the population)?''
\end{quote}
A paragraph has now been added to Section 4.1 comparing allocations that maximise survival, maximise coverage, and minimise response times.

@@

\item Reviewer 2 wrote:
\begin{quote}
``Is the survival function the same for $K_a$ patients regardless of which vehicle visits them? What is the rationale for this? It seems that primary vehicles would be better prepared to deal with complications, etc.''
\end{quote}
This is an interesting suggestion. Currently we model the survival function to be independent of the vehicle type, that is $s_k$ is the same regardless if the vehicle that arrived is a primary or secondary vehicle. As you note, this may not be realistic, and we may want to be able to model situations where there are different survival functions for each vehicle type. However, this becomes more complicated when both vehicles arrive, but at different times - there will be an interaction between the functions that requires more investigation and is beyond the scope of this study. These scenarios are interesting, and we have added a sentence in the paper explaining this as further work.

\item Reviewer 2 wrote:
\begin{quote}
``Have you given any consideration to using an exact method to solve the optimisation problem solved in Algorithm 1? Is there any alternative in the literature?''
\end{quote}
We did not consider using an exact method to solve this as the were so many non-linear elements to the problem, namely the objective function and the utilisation considerations. An exact method would be beyond the scope of this paper, which focusses on the non-linear objective function formulation, and simulation modelling.

\item Reviewer 2 wrote:
\begin{quote}
``P11 L43 – What is the rationale of using a triangular distribution for travel times? In general, what is the basis of all the travel time assumptions presented in this section?''
\end{quote}
A reference has now been given justifying the use of triangular distributions when little information is known. All other travel time information is given in the Appendix, where we state that travel times are derived using Google Maps API, and traffic delays found using the TomTom API.

@@

\item Reviewer 2 wrote:
\begin{quote}
``P12 5.4 – For patients that might not need to travel to a hospital, would they still need a primary vehicle after a secondary vehicle tends to them? If not the logic would not be separate entirely.''
\end{quote}
This is a very good point, and is similar to the situation when a primary vehicle reaches the patient before the secondary vehicle. Ideally in these cases the later vehicle would turn around and be free to pick up another patient. However one assumption of the model is that en-route communications are difficult and do not occur. We have now emphasised this in the text.

\item Reviewer 2 wrote:
\begin{quote}
``It seems there are two conflicting methods here, the optmisiation with the EA, which uses some averages, etc. And a more comprehensive DES simulation; however the latter is not used for the optimisation. Did you consider simulation optimisation or simheuristics to address this?''
\end{quote}
We did not consider using the simulation as part of the optimisation process, and is beyong the scope of the paper. However, we think this is a great idea, and have discussed this as future work in the paper, as a tool to give greater consideration to the variability in the system.

\item Reviewer 2 wrote:
\begin{quote}
``P15 Paragraph starting L36. Can you give more detail on the modelling of the call arrival rates? For rare specialties, and given the low usage of ambulances in Indonesia, you might not have enough data to give an accurate representation of actual demand. Especially, for future demand where visibility is increased, how will this change for example among richer/poorer parts of the city? I acknowledge it is impossible to include everything in your experiments, but how do you expect these limitations might affect your results?''
\end{quote}
The call arrival rates are taken directly from provided data, and reported in the Appendix. You make a good point regarding the pick-up locations with particularly low arrival rates - this is not a problem for \textbf{D13} as this is what was observed, but might result in artificially low arrival rates when multiplying up by demand. This is now discussed in the paper.

\item Reviewer 2 wrote:
\begin{quote}
``Could you conduct any validation of your results with real data from eg. Ambulance response times with the baseline allocation?''
\end{quote}
Thorough verification of the models is done by the automated code testing suits provided in the code base, a paragraph on this is now included under Section 5.5. Validation of this model was difficult, and relied on the expert opinion of our collaborators ar 118 and 119. This is now discussed in Section 6.

### 2024-07-31 tex/main.tex
Complete first draft of response and review

This is the method used to calculate the expected values of $B_{pa}$, $C_{py}$,
$D_{ya}$ and $F_{pa}$.
Furthermore it is assumed that travel times for each segment of the route
(from vehicle location to patient, from patient to hospital, and from hospital
back the the vehicle's location) follow a Triangular distribution around the
calculated expected travel time, between 75\% and 125\% of the value.

@@

\subsection{Combined Model}
Combining the logic presented in Sections~\ref{sec:simulation_primary}
and~\ref{sec:simulation_secondary} in a sequential manner, with the output of
one determining the input of the other, gives the overall simulation logic for
simulating both primary and secondary vehicle types. It is used to quantify
the effectiveness of a given allocation $(Z_a, \tilde{Z}_a)$ for all
$a \in \mathcal{A}$, by recording several useful KPIs.
In this work, the KPIs of interest are: the average primary vehicle
utilisation, the average secondary vehicle utilisation, the mean response
time, the percentage of abandoned calls (that is a measure of the primary
vehicle unavailability), and the expected overall survival. Survival is
modelled using a combination of survival function curves and hard cut-offs,
and is described in Section~\ref{sec:survival}.
The source code for the combined simulation, is available at \citet{zenodo_sim} and
development took place at \url{https://github.com/MarkTuson/ambulance_simulation}.

@@

All models are parameterised using demand data that covers all 261
neighbourhoods from 1 January to 31 December 2019, before the COVID-19
pandemic. Data from 2020-2021 was naturally heavily skewed by the pandemic,
with significantly lower demand, and so was not considered as representative
of a typical period for forecasting future needs, hence the decision to use
the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters} provides further details on the
parameterisation of the model, including how travel time estimates were
obtained.

@@

We consider four different demand scenarios, developed by considering the
result of our cross-sectional study of patients attending EDs in Jakarta
\citep{BriceSyaribahNoor2022Esui}. As part of the survey, randomly selected
patients arriving at each emergency department were asked whether they had
used an ambulance to attend, and reasons for not using an ambulance. For those
patients who were categorised by the medical staff as emergency and for whom
therefore an ambulance might have seemed a sensible option, the responses are
given in Table~\ref{table:survey_results}.

@@

Guided by our findings in \citet{BriceSyaribahNoor2022Esui}, recent changes
mean that the 119 services is now free to use for Jakarta residents, so
certainly the highest demand (\textbf{D45}) scenario is plausible in the near
future. After discussions with our partners, it was agreed that for each of
the scenarios we should assume that non-emergency demand (speciality B)
remains unchanged.

>>>>

This is the method used to calculate the expected values of $B_{pa}$, $C_{py}$,
$D_{ya}$ and $F_{pa}$.
Furthermore it is assumed that travel times for each segment of the route
(from vehicle location to patient, from patient to hospital, and from hospital
back the the vehicle's location) follow a Triangular distribution around the
calculated expected travel time, between 75\% and 125\% of the value.
\textcolor{red}{A Triangular distribution is a useful distribution to use when
not much is known about the travel times except some basic summary statistics
\citep{robinson2014simulation}, in this case a likely range of values.}

@@

\subsection{Combined Model}\label{sec:combined_model}
Combining the logic presented in Sections~\ref{sec:simulation_primary}
and~\ref{sec:simulation_secondary} in a sequential manner, with the output of
one determining the input of the other, gives the overall simulation logic for
simulating both primary and secondary vehicle types. It is used to quantify
the effectiveness of a given allocation $(Z_a, \tilde{Z}_a)$ for all
$a \in \mathcal{A}$, by recording several useful KPIs.
In this work, the KPIs of interest are: the average primary vehicle
utilisation, the average secondary vehicle utilisation, the mean response
time, the percentage of abandoned calls (that is a measure of the primary
vehicle unavailability), and the expected overall survival. Survival is
modelled using a combination of survival function curves and hard cut-offs,
and is described in Section~\ref{sec:survival}.

\textcolor{red}{
Verification of a model is ensuring that the conceptual model is implemented
correctly and without error, while validation of a model ensures that the
conceptual model accurately represents the real system to be modelled
\citep{robinson2014simulation}. Verifying the simulation code, alongside
verifying the optimisation and MESLMHPHF objective function code, is done
through automated code testing \citep{percival2014test, palmer2023implementing}.
This includes unit testing to ensure that each sub-component works are requires,
and is bug and error free; end to end testing, ensuring that the components
interact together as intended; and extreme value testing ensuring that changes
in input parameters result in sensible and expected changes in output results
and behaviours. The testing suits are included in the open code repositories
provided.
}

The source code for the combined simulation, is available at \citet{zenodo_sim}
and development took place at \url{https://github.com/MarkTuson/ambulance_simulation}.

@@

All models are parameterised using demand data that covers all 261
neighbourhoods from 1 January to 31 December 2019, before the COVID-19
pandemic. Data from 2020-2021 was naturally heavily skewed by the pandemic,
with significantly lower demand, and so was not considered as representative
of a typical period for forecasting future needs, hence the decision to use
the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters}
provides further details on the parameterisation of the model, including how
travel time estimates were obtained.
\textcolor{red}{Verification of the models were discussed in
Section~\ref{sec:combined_model}, however parameter validation for this
particular case study is difficult, with all primary data available to us being
used to parametrise the model, still leaving many aspects to be determined by
expert opinion. This expert opinion however was vital in giving a basic
validation of the model, with staff from Ambulans 118 and the 119 Emergency
Ambulance Service in Jakarta approving model results parametrised from their
provided data and expert opinion.}

@@

\textcolor{red}{In other cities with more detailed data collection, statistical
models can be used to predict future spacial demand across a city, as in
\cite{nicoletta2022bayesian}. In the case of Jakarta, collected call data is
not detailed enough, but we do}
consider four different demand scenarios, developed by considering the
result of our cross-sectional study of patients attending EDs in Jakarta
\citep{BriceSyaribahNoor2022Esui}. As part of the survey, randomly selected
patients arriving at each emergency department were asked whether they had
used an ambulance to attend, and reasons for not using an ambulance. For those
patients who were categorised by the medical staff as emergency and for whom
therefore an ambulance might have seemed a sensible option, the responses are
given in Table~\ref{table:survey_results}.

@@

Guided by our findings in \citet{BriceSyaribahNoor2022Esui}, recent changes
mean that the 119 services is now free to use for Jakarta residents, so
certainly the highest demand (\textbf{D45}) scenario is plausible in the near
future. After discussions with our partners, it was agreed that for each of
the scenarios we should assume that non-emergency demand (speciality B)
remains unchanged.
\textcolor{red}{Therefore, in each scenario, we multiply the call demand for
each municipality, $\lambda_{pk}$ by $\frac{13\%}{13\%}=1$,
$\frac{19\%}{13\%}=1.46$, $\frac{34\%}{13\%}=2.62$, and
$\frac{45\%}{13\%}=3.46$ respectively, for all $k \in \mathcal{K}_A$,
$p \in \mathcal{P}$. Note that for some pick-up locations $p$, $\lambda_{pk}$
is very small, possibly zero, which is realistic in scenario \textbf{D13}, as
that is what is observed, but may result in artificially low demand in those
locations when using the multiplication method to scale up demand.}
